---
title: "Sentiment Analysis of Star Wars (4-6) Movie Scripts"
output:
  html_document:
    toc: true
    toc_float: true
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

# Dependencies

```{r}
library(tidyverse)
library(tidytext)
```

# Reading in clean data

```{r}
sw_scripts <- read_csv("../clean_data/original_sw_trilogy.csv")
```

# Tokenize and remove stop words

I will use single words as my token as I am interested in the sentiments of words.
```{r}
sw_tokens <- sw_scripts %>%
  unnest_tokens(
    word,
    dialogue
  ) %>%
  anti_join(stop_words)

sw_tokens
```

# Check which sentiment lexicon categorizes most words

Before analysing the sentiment of the text, I want to check which lexicon is able to categorise/rate most words.

Available lexicons in `tidytext::get_gentiments()`:

* __Bing__
  * 2 categories: positive or negative
* __AFINN__
  * 11 ratings: integer between -5 (negative) and +5 (positive)
* __Loughran__
  * 6 ratings: negative, positive, litigious, uncertainty, constraining, or superfluous 
* __NRC__
  * 10 categories: 8 basic emotions (anger, fear, anticipation, trust, surprise, sadness, joy, or disgust) and 2 sentiments (negative or positive)
  
```{r}
unique_words <- sw_tokens %>% distinct(word)

# available lexicons in tidytext::get_sentiments()
lexicons <- c("bing", "afinn", "loughran", "nrc")

# create list of joined datasets with available lexicons
nested_df <- lexicons %>%
  map(~left_join(unique_words, get_sentiments(.), by = "word"))

# attach lexicon names to list
names(nested_df) <- lexicons 


for (lexicon in lexicons){

  # 2nd element is sentiment category or rating
  sentiments <- nested_df[[lexicon]][[2]]
  
  # count all values without attached sentiment
  missing <- sum(is.na(sentiments)) 
  
  print(str_glue("{lexicon}: {missing} uncategorised words"))

}
```

Seems like the NRC lexicon is able to categorise most words, so I will use it for my sentiment analysis.

# Which character expresses the most sentiment

## Set theme for all plots

```{r}
theme_set(theme_minimal() + 
            theme(
              strip.text = element_text(size = 10, face = "bold"),
              axis.text = element_text(size = 8, face = "bold"),
              axis.title = element_text(size = 12, face = "bold"),
              plot.title = element_text(size = 15, face = "bold")
              )
          )
```

## Create dataset with all NRC-categorised script words

```{r}
sw_sentiments <- sw_tokens %>% 
  inner_join(get_sentiments("nrc"), by = "word")
```

## Sentiment words spoken by characters

```{r}
# need to account for how much characters speak in general
# select top 20 speaking characters and calculate percentage?


sw_sentiments %>% 
  filter(!(sentiment %in% c("positive", "negative"))) %>% 
  group_by(character, sentiment) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  arrange(desc(count)) %>% 
  group_by(sentiment) %>% 
  slice_max(count, n = 5) %>% 
  mutate(is_max_count = count == max(count),
         character = recode(character, Threepio = "C-3PO")) %>% 
  ggplot(aes(
    x = reorder_within(character, -count, sentiment), 
    y = count,
    fill = is_max_count
    )) +
  geom_col() +
  scale_fill_manual(
    values = c(`TRUE` = "#4074AD", `FALSE` = "grey70"), guide = F) +
  labs(
    x = NULL,
    y = "Number of words",
    title = "Emotions expressed by Star Wars characters (Ep 4-6)",
    subtitle = "Top 5 characters using NRC sentiment words"
  ) +
  facet_wrap(~sentiment, scales = "free_x", dir = "v") +
  scale_x_reordered()
```


```{r, echo = F, include = F}
ggsave(filename="../plots/sentiment_words.png", 
       device="png", 
       height=5, 
       width=8, 
       units="in", 
       dpi=500)
```

# Sentiment arcs

```{r}
sw_tokens %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(episode) %>% 
  mutate(
    word_n = row_number()
  ) %>%
  mutate(
    story_position = word_n/max(word_n) # all books on scale from 0 to 1
  ) %>% 
  mutate(episode = as.character(episode),
         episode = recode(episode, 
                          "4" = "Episode IV: A New Hope",
                          "5" = "Episode V: The Empire Strikes Back",
                          "6" = "Episode VI: Return of the Jedi")) %>% 
  ggplot() +
  aes(x = story_position, y = value, colour = episode) +
  geom_smooth(se = FALSE) +
  facet_wrap(~episode, ncol = 1) +
  coord_cartesian(ylim = c(-3, 3)) +
  scale_colour_manual(values = c("#248AD2",
                               "#6A8FBB",
                               "#4074AD"),
                    guide = "none") +
  labs(x = "Story progression",
       y = "Sentiment rating") +
  labs(
    x = NULL,
    y = "Sentiment rating",
    title = "Sentiment arcs of original Star Wars trilogy",
    subtitle = "Sentiment ratings from AFINN lexicon"
  )

```

```{r, echo = F, include = F}
ggsave(filename="../plots/sentiment_arcs.png", 
       device="png", 
       height=5, 
       width=8, 
       units="in", 
       dpi=500)
```
